---
title: "Study 2"
output:
  html_document:
    toc: true
    theme: default
---

## (1) Description of Data Files

```{r include=FALSE}
options(width=100)

library(data.table)
library(pequod)
library(psych)
library(lm.beta)
library(rtf)
library(RCurl)

# load data
csv <- getURL("https://raw.githubusercontent.com/abrowman/psm-jesp2017/master/Study%202/SocMobility_Study2.csv")
mydata <- read.csv(textConnection(csv))
```

The posted data for this study consists of 1 compiled, deidentified data file (total *N* recruited = `r nrow(mydata) + 2`):  

* 2 responses excluded prior to compiling this deidentified file because (1) 1 participant completed the study twice, so their second response was removed, and (2) 1 participant started the study twice but they did not continue beyond the consent page the 1st time, so their 2nd response was retained\
* `r nrow(subset(mydata, year>4 | is.na(year)))` participants excluded because they were not undergraduate students\
* `r nrow(subset(mydata, (year<=4 & !is.na(year)) & (!is.na(dircheck1) | !is.na(dircheck2))))` participants were excluded because they failed at least 1 attention check\
      \

```{r include=FALSE}
mydata <- subset(mydata, mydata$year<=4) # remove non-undergraduates
mydata <- subset(mydata, is.na(dircheck1) & is.na(dircheck2)) # removed those who failed at least 1 attention check
attach(mydata)

# determine condition and set contrast codes to weak = -1, strong = 1
cond <- ifelse(!is.na(entity1),-1,1)
cond_c <- scale(cond,scale=F,center=T)
cond <- factor(ifelse(cond==-1, "weak", "strong"))
options(contrasts=c("contr.sum","contr.poly"))
contrasts(cond) = c(-1, 1)
contrasts(cond)

# calculate scores
psm <- (ltss1+(8-ltss2r)+ltss3+(8-ltss4r))/4
lti <- ((8-lti1r)+lti2+(8-lti3r))/3
ldr <- 11-ladder # Qualtric coded ladder in reverse
acc <- (correctwords/totalwords)*100

# center ses predictors
ladder_c <- scale(ldr,scale=F,center=T)
income_c <- scale(income,scale=F,center=T)
lti_c <- scale(lti,scale=F,center=T)
```

## (2) Descriptive Statistics

**Total *N*:** `r nrow(mydata)`

***N* in each condition:**
```{r echo=FALSE}
table(cond, useNA="always")
```

**Percentage by gender:**
```{r echo=FALSE}
Gender <- ifelse(gender==1, "male", ifelse(gender==2, "female", NA))
table(Gender, useNA="always")

gender_c <- scale(ifelse(Gender=="male", 1, ifelse(Gender=="female", -1, NA)), scale=F,center=T)
```

**Percentage by college class:**
```{r echo=FALSE}
class <- ifelse(year==1, "freshman", ifelse(year==2, "sophomore", ifelse(year==3, "junior", ifelse(year==4, "senior", 4))))
round((table(class, useNA="always")/nrow(mydata))*100, digits = 1)
```

**Percentage by college type:**
```{r echo=FALSE}
table(collegetype, useNA="always")
```

***M*s and *SD*s:**
```{r echo=FALSE}
describe(data.frame(psm,ldr,income,lti))[c(2,3,4,8,9)]
cor.test(income,ldr)
```

**Alphas:**

* **Perceptions of socioeconomic mobility (psm; manipulation check):** `r round(alpha(data.frame(ltss1,(8-ltss2r),ltss3,(8-ltss4r)))$total$raw_alpha, digits=2)`  
* **Lay theories of intelligence (lti):** `r round(alpha(data.frame((8-lti1r),lti2,(8-lti3r)))$total$raw_alpha, digits=2)` 


## (3) Data Screening

### Random assignment checks

**(a) Unequal variance test, *t*-test, and *M*s and *SD*s for: subjective SES**
```{r echo=FALSE}
var.test(ldr~cond)
t.test(ldr~cond, var.equal=TRUE)
describeBy(ldr,cond)
```

**(b) Unequal variance test, *t*-test, and *M*s and *SD*s for: income**
```{r echo=FALSE}
var.test(income~cond)
t.test(income~cond, var.equal=TRUE)
describeBy(income,cond)
```

### Manipulation checks

**(a) Unequal variance test, *t*-test, *M*s and *SD*s, and Cohen's *D* for: Perceptions of socioeconomic mobility**
```{r echo=FALSE}
var.test(psm ~ cond)
t.test(psm ~ cond, var.equal=TRUE)
describeBy(psm,cond)
```

**Cohen's *d*:** `r round((describeBy(psm, cond)$malleable$mean-describeBy(psm, cond)$fixed$mean)/(sqrt((describeBy(psm, cond)$malleable$sd^2+describeBy(psm, cond)$fixed$sd^2)/2)), digits=2)`  
  

**(b) Perceptions of mobility by condition and SES**
```{r echo=FALSE}
summary(lm.beta(lm(psm ~ cond_c * income_c)))
```

**95% confidence intervals for unstandardized and standardized regression coefficients:**
```{r echo=FALSE}
psm.regress <- lm(psm ~ cond_c * income_c)
psm.regress.ci <- data.frame(round(confint(psm.regress, level = 0.95), digits=2))
colnames(psm.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
psm.regress.ci
```


**(c) Unequal variance test, *t*-test, and *M*s and *SD*s for: LTI**
```{r echo=FALSE}
var.test(lti ~ cond)
t.test(lti ~ cond, var.equal=TRUE)
describeBy(lti,cond)
```

### Correcting negative skew in anagram score

The raw anagram scores were highly negatively skewed (skew = `r round(describe(acc)$skew, digits=2)`). As such, they were subjected to a logarithmic transformation before being submitted to our analyses (see Ratcliff, 1993). This produced an acceptable level of skew (`r round(describe(-log10(101-acc)+2.79)$skew, digits=2)`), whereas using an inverse transformation did not (`r round(describe((1/(101-acc))+.98)$skew, digits=2)`).

```{r include=FALSE}
t.acc <- -log10(101-acc)+2.79 # log transformation
```

## (4) Main Analyses

### Analyses with income and without covariates

**Correlation with persistence**
```{r echo=FALSE}
cor.test(t.acc, income)
```

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * income_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
inc.regress <- lm(t.acc ~ cond_c * income_c)
inc.regress.ci <- data.frame(round(confint(inc.regress, level = 0.95), digits=2))
colnames(inc.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
inc.regress.ci
```

**Simple slopes tests:**
```{r echo=FALSE}
inc.lmres <- lmres(t.acc ~ cond_c * income_c, mydata)
summary(inc.ss <- simpleSlope(inc.lmres, pred="cond_c", mod1="income_c"))
```

**CIs among low-income students:** `r round(inc.ss$simple_slope[1,1] + c(-1,1)*inc.ss$simple_slope[1,2]*qt(0.975, inc.ss$Df), digits=2)`  

**CIs among high-income students:** `r round(inc.ss$simple_slope[2,1] + c(-1,1)*inc.ss$simple_slope[2,2]*qt(0.975, inc.ss$Df), digits=2)`

**Plot (with untransformed data):**  
```{r echo=FALSE, warning=FALSE}
inc.nontrans.lmres <- lmres(acc ~ cond_c * income_c, mydata)
inc.nontrans.ss <- simpleSlope(inc.nontrans.lmres, pred="cond_c", mod1="income_c")
inc.plot.points <- data.frame(c("Fixed","Malleable","Fixed","Malleable"), c("aLow-SES","aLow-SES","bHigh-SES","bHigh-SES"), c(inc.nontrans.ss$Points[1,1], inc.nontrans.ss$Points[1,2], inc.nontrans.ss$Points[2,1], inc.nontrans.ss$Points[2,2]), c(describe(acc)$se, describe(acc)$se, describe(acc)$se, describe(acc)$se))
colnames(inc.plot.points) <- c("psm","ses","score","se")
limits <- aes(ymax=score+se, ymin=score-se)

lo.title <- expression(paste("Lower-SES students (-1 ", italic("SD"), ")"))
hi.title <- expression(paste("Higher-SES students (+1 ", italic("SD"), ")"))

ggplot(inc.plot.points, aes(x=ses, y=score, fill=psm)) +
  geom_bar(stat="identity",position=position_dodge(width=.6), colour="black", width=.5) +
  scale_fill_manual(values=c("gray", "white"),
                    breaks=c("Fixed", "Malleable"),
                    labels=c("  Weak mobility beliefs condition     ", "  Strong mobility beliefs condition     ")) +
  geom_errorbar(limits, position=position_dodge(width=.6), width=.1) +
  geom_hline(yintercept=0) +
  theme_bw(base_family = "Arial", base_size = 16) + # font and font size
  theme(legend.position = "bottom") + # legend position
  theme(legend.title = element_blank()) + # remove legend title
  theme(axis.title.x = element_blank()) + # move x-axis label lower
  theme(axis.title.y = element_text(vjust=1.3)) + # move y-axis label left
  theme(panel.grid.minor = element_blank()) + # no grids
  theme(panel.grid.major = element_blank()) + # no borders
  theme(panel.border = element_blank()) + # removes default grey axis lines
  theme(axis.line = element_line(color = 'black')) + # put axis lines back, but in black
  coord_cartesian(ylim=c(50, 100)) + # y-axis range
  scale_x_discrete("", labels = c("aLow-SES"=lo.title, "bHigh-SES"=hi.title)) + # x-axis titles
  scale_y_continuous("Anagram scores (%)", breaks = seq(50, 100, by = 5)) # y-axis titles and range
ggsave('Figure 2.png', width=9, height=6, unit='in', dpi=300)
```


### Analyses with subjective SES and without covariates

**Correlation with persistence**
```{r echo=FALSE}
cor.test(t.acc, ladder)
```

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * ladder_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
ldr.regress <- lm(t.acc ~ cond_c * ladder_c)
ldr.regress.ci <- data.frame(round(confint(ldr.regress, level = 0.95), digits=2))
colnames(ldr.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
ldr.regress.ci
```

**Simple slopes tests:**
```{r echo=FALSE}
ldr.lmres <- lmres(t.acc ~ cond_c * ladder_c, mydata)
summary(ldr.ss <- simpleSlope(ldr.lmres, pred="cond_c", mod1="ladder_c"))
```

**CIs among low subjective SES students:** `r round(ldr.ss$simple_slope[1,1] + c(-1,1)*ldr.ss$simple_slope[1,2]*qt(0.975, ldr.ss$Df), digits=2)`  

**CIs among high subjective SES students:** `r round(ldr.ss$simple_slope[2,1] + c(-1,1)*ldr.ss$simple_slope[2,2]*qt(0.975, ldr.ss$Df), digits=2)`


### Analyses with income and covariates

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * income_c + lti_c * income_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
inc.cov.regress <- lm(t.acc ~ cond_c * income_c + lti_c * income_c)
inc.cov.regress.ci <- data.frame(round(confint(inc.cov.regress, level = 0.95), digits=2))
colnames(inc.cov.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
inc.cov.regress.ci
```

**Simple slopes tests:**
```{r echo=FALSE}
inc.cov.lmres <- lmres(t.acc ~ cond_c * income_c + lti_c * income_c, mydata)
summary(inc.cov.ss <- simpleSlope(inc.cov.lmres, pred="cond_c", mod1="income_c"))
```

**CIs among low-income students:** `r round(inc.cov.ss$simple_slope[1,1] + c(-1,1)*inc.cov.ss$simple_slope[1,2]*qt(0.975, inc.cov.ss$Df), digits=2)`  

**CIs among high-income students:** `r round(inc.cov.ss$simple_slope[2,1] + c(-1,1)*inc.cov.ss$simple_slope[2,2]*qt(0.975, inc.cov.ss$Df), digits=2)`


### Analyses with subjective SES and covariates

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * ladder_c + lti_c * ladder_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
ldr.cov.regress <- lm(t.acc ~ cond_c * ladder_c + lti_c * ladder_c)
ldr.cov.regress.ci <- data.frame(round(confint(ldr.cov.regress, level = 0.95), digits=2))
colnames(ldr.cov.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
ldr.cov.regress.ci
```

**Simple slopes tests:**
```{r echo=FALSE}
ldr.cov.lmres <- lmres(t.acc ~ cond_c * ladder_c + lti_c * ladder_c, mydata)
summary(ldr.cov.ss <- simpleSlope(ldr.cov.lmres, pred="cond_c", mod1="ladder_c"))
```

**CIs among low subjective SES students:** `r round(ldr.cov.ss$simple_slope[1,1] + c(-1,1)*ldr.cov.ss$simple_slope[1,2]*qt(0.975, ldr.cov.ss$Df), digits=2)`  

**CIs among high subjective SES students:** `r round(ldr.cov.ss$simple_slope[2,1] + c(-1,1)*ldr.cov.ss$simple_slope[2,2]*qt(0.975, ldr.cov.ss$Df), digits=2)`


### Analyses with gender

**T-test:**
```{r echo=FALSE}
var.test(t.acc ~ Gender)
t.test(t.acc ~ Gender, var.equal=T)
describeBy(t.acc, Gender)
```

#### Analyses with income

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * income_c * gender_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
gender.inc.regress <- lm(t.acc ~ cond_c * income_c * gender_c)
gender.inc.regress.ci <- data.frame(round(confint(gender.inc.regress, level = 0.95), digits=2))
colnames(gender.inc.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
gender.inc.regress.ci
```


#### Analyses with subjective SES

**Omnibus test:**
```{r echo=FALSE}
summary(lm.beta(lm(t.acc ~ cond_c * ladder_c * gender_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
gender.ldr.regress <- lm(t.acc ~ cond_c * ladder_c * gender_c)
gender.ldr.regress.ci <- data.frame(round(confint(gender.ldr.regress, level = 0.95), digits=2))
colnames(gender.ldr.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
gender.ldr.regress.ci
```

**Simple effects:**
```{r echo=FALSE}
gender.ldr.lmres <- lmres(t.acc ~ cond_c * ladder_c * gender_c, mydata)
summary(gender.ldr.ss <- simpleSlope(gender.ldr.lmres, pred="cond_c", mod1="gender_c", mod2="ladder_c"))
```

**CIs among female low subjective SES students:** `r round(gender.ldr.ss$simple_slope[1,1] + c(-1,1)*gender.ldr.ss$simple_slope[1,2]*qt(0.975, gender.ldr.ss$Df), digits=2)`  

**CIs among female high subjective SES students:** `r round(gender.ldr.ss$simple_slope[2,1] + c(-1,1)*gender.ldr.ss$simple_slope[2,2]*qt(0.975, gender.ldr.ss$Df), digits=2)`

**CIs among male low subjective SES students:** `r round(gender.ldr.ss$simple_slope[3,1] + c(-1,1)*gender.ldr.ss$simple_slope[3,2]*qt(0.975, gender.ldr.ss$Df), digits=2)`  

**CIs among male high subjective SES students:** `r round(gender.ldr.ss$simple_slope[4,1] + c(-1,1)*gender.ldr.ss$simple_slope[4,2]*qt(0.975, gender.ldr.ss$Df), digits=2)`


```{r include=FALSE, warning=FALSE}
## Create and export tables ##

# full model sections
regtable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), round(coef(summary(modellist[[i]])), digits=3))
    assign(paste("ci", i, sep = ""), paste(round(get(paste("b", i, sep=""))[, 1], digits=2), " [", round(confint(modellist[[i]], level = 0.95)[,1], digits=2), ", ", round(confint(modellist[[i]], level = 0.95)[,2], digits=2), "]", sep = ""))
    assign(paste("t", i, sep = ""), round(summary(modellist[[i]])$coefficients[,3], digits=2))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$df[2],each=1+length(labels(modellist[[1]]))))
    assign(paste("p", i, sep = ""), round(summary(modellist[[i]])$coefficients[,4], digits=3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(all.vars(summary(modellist[[i]])$call)[1], ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rnames <- c("Intercept")
  for (j in 1:length(labels(modellist[[1]]))) {
    rnames <- append(rnames, labels(modellist[[1]])[j])
  }
  rownames(complete) <- rnames
  complete <- complete[-1,]
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=8)  # this can be an .rtf or a .doc
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models1 <- list(inc.regress)
regtable(models1, "Table 2a (omnibus)")
models2 <- list(inc.cov.regress)
regtable(models2, "Table 2b (omnibus)")
models3 <- list(ldr.regress)
regtable(models3, "Table S3a (omnibus)")
models4 <- list(ldr.cov.regress)
regtable(models4, "Table S3b (omnibus)")
models5 <- list(gender.inc.regress, gender.ldr.regress)
regtable(models5, "Table S4 (gender)")

# simple slope sections
sstable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), round(summary(modellist[[i]])$simple_slope, digits=3))
    assign(paste("ci", i, sep = ""), paste(round(get(paste("b", i, sep=""))[, 1], digits=2), " [", c(round(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), digits=2)[1], round(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[i]]$Df), digits=2)[1]), ", ", c(round(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), digits=2)[2], round(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[1]]$Df), digits=2)[2]), "]", sep = ""))
    assign(paste("t", i, sep = ""), round(get(paste("b", i, sep=""))[,3], digits=2))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$Df,each=2))
    assign(paste("p", i, sep = ""), round(get(paste("b", i, sep=""))[,4], digits=3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(modellist[[i]]$nomY, ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c(row.names(summary(modellist[[i]])$Points)[1], row.names(summary(modellist[[i]])$Points)[2])
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=12)
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models1.ss <- list(inc.ss)
sstable(models1.ss, "Table 2 (simple slopes)")
models2.ss <- list(ldr.ss)
sstable(models2.ss, "Table S3 (simple slopes)")
models3.ss <- list(inc.cov.ss, ldr.cov.ss)
sstable(models3.ss, "Table S4 (simple slopes)")
```
