---
title: "Study 1"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: default
---

## (1) Description of Data Files

```{r include=FALSE}
options(width=100)

library(plyr)
library(pequod)
library(psych)
library(lm.beta)
library(ppcor)
library(data.table)
library(rtf)
library(RCurl)
library(mediation)

# load data
csv <- getURL("https://raw.githubusercontent.com/abrowman/psm-jesp2017/master/Study%201/SocMobility_Study1.csv")
mydata <- read.csv(textConnection(csv))

# recode values
mydata$gender <- gsub("F", "f", mydata$gender)
mydata$gender <- gsub("famale", "f", mydata$gender)
mydata$gender <- gsub("female ", "f", mydata$gender)
mydata$gender <- gsub("female", "f", mydata$gender)
mydata$gender <- gsub("Female ", "f", mydata$gender)
mydata$gender <- gsub("Female", "f", mydata$gender)
mydata$gender <- gsub("famale", "f", mydata$gender)
mydata$gender <- gsub("G", "f", mydata$gender)
mydata$gender <- gsub("girl", "f", mydata$gender)
mydata$gender <- gsub("mother", "f", mydata$gender)
mydata$gender <- gsub("mALE", "m", mydata$gender)
mydata$gender <- gsub("male ", "m", mydata$gender)
mydata$gender <- gsub("male", "m", mydata$gender)
mydata$gender <- gsub("Male ", "m", mydata$gender)
mydata$gender <- gsub("Male", "m", mydata$gender)
mydata$gender <- gsub("MALE", "m", mydata$gender)
mydata$gender <- gsub("man", "m", mydata$gender)
mydata$gender <- gsub("Man", "m", mydata$gender)
mydata$gender <- gsub("men", "m", mydata$gender)
mydata$gender <- factor(ifelse(mydata$gender=="m" | mydata$gender=="f", mydata$gender, NA))
mydata$gender <- factor(ifelse(mydata$gender=="m", "male", ifelse(mydata$gender=="f", "female", NA)))
mydata$age <- gsub("1/2", "", mydata$age)
mydata$age <- gsub("going on 15", "14", mydata$age)
mydata$age <- as.numeric(gsub("34", "", mydata$age))
attach(mydata)
options(contrasts=c("contr.sum","contr.poly"))

# score measures
psm <- ((8-ltss1r)+(8-ltss2r)+(8-ltss3r)+ltss4+ltss5+ltss6)/6
psm_c <- scale(psm, scale=F,center=T)
lti <- ((8-lti1r)+(8-lti2r)+lti3+lti4+(8-lti5r)+lti6)/6
lti_c <- scale(lti, scale=F,center=T)
persist <- 8-((diff.imposs1+diff.imposs2+diff.imposs3+diff.imposs4)/4)
persist_c <- scale(persist, scale=F,center=T)
```

The posted data for this study consists of 1 compiled, deidentified data file (total *N* recruited = `r nrow(mydata)`):

* No data were excluded

## (2) Descriptive statistics

**Total *N*:** `r nrow(mydata)`

**Percentage by gender:**
```{r echo=FALSE}
round(table(gender, useNA="always"), digits=1)

gender_c <- scale(ifelse(gender=="male", 1, ifelse(gender=="female", -1, NA)), scale=F,center=T)
```

**Percentage by grade:**
```{r echo=FALSE}
Grade <- ifelse(grade==1, "9th grade", ifelse(grade==2, "10th grade", ifelse(grade==3, "11th grade", ifelse(grade==4, "12th grade", NA))))
round((table(Grade, useNA="always")/nrow(mydata))*100, digits=1)
```

***M*s and *SD*s:**
```{r echo=FALSE}
describe(data.frame(age, psm, persist, gpa, lti))[c(2,3,4,8,9)]
```

**Alphas:**  

  * **Perceptions of socioeconomic mobility:** `r round(alpha(data.frame((8-ltss1r),(8-ltss2r),(8-ltss3r),ltss4,ltss5,ltss6))$total$raw_alpha, digits=2)`
  * **Inclination to persist academically:** `r round(alpha(data.frame(diff.imposs1,diff.imposs2,diff.imposs3,diff.imposs4))$total$raw_alpha, digits=2)`
  * **Lay theories of intelligence:** `r round(alpha(data.frame((8-lti1r),(8-lti2r),lti3,lti4,(8-lti5r),lti6))$total$raw_alpha, digits=2)`


## (3) Manipulations checks

**Perceptions of socioeconomic mobility:**
```{r echo=FALSE}
summary(aov(psm ~ cond))
```

**Academic persistence:**
```{r echo=FALSE}
summary(aov(persist ~ cond))
```

**Academic performance:**
```{r echo=FALSE}
summary(aov(gpa ~ cond))
```


## (4) Main analyses

### Bivariate correlations
```{r echo=FALSE}
corlist <- data.frame(psm, persist, gpa, lti)
cortable <- data.frame(c(corr.test(as.matrix(corlist))$n)[c(2,3,4,7,8,12)]-2, round(corr.test(as.matrix(corlist))$ci[,1:3], digits=2), round(corr.test(as.matrix(corlist))$ci[,4], digits=4))
colnames(cortable) <- c("df","lower","r","upper","p")
cortable
```

### Partial correlations, controlling for LTI

#### Perceptions of socioeconomic mobility and persistence:
```{r echo=FALSE}
cordat1 <- subset(data.frame(psm, persist, lti), !is.na(psm) & !is.na(persist) & !is.na(lti))
pcorr.psm.persist <- pcor.test(cordat1$psm, cordat1$persist, cordat1$lti)
pcorr.psm.persist
```

```{r include=FALSE}
corr.ci <- function(alpha, corr, s, n) {
  z <- qnorm(1 - alpha/2)
  se <- sqrt(1/((n - s - 3)))
  zr <- log((1 + corr)/(1 - corr))/2
  LL0 <- zr - z*se
  UL0 <- zr + z*se
  LL <- (exp(2*LL0) - 1)/(exp(2*LL0) + 1)
  UL <- (exp(2*UL0) - 1)/(exp(2*UL0) + 1)
  CI <- c(LL, UL)
  return(round(CI, digits=2))
}
```

**Confidence intervals:** `r corr.ci(alpha = .05, corr = pcorr.psm.persist$estimate, s = 1, n = pcorr.psm.persist$n)`

#### Perceptions of socioeconomic mobility and GPA:
```{r echo=FALSE}
cordat2 <- subset(data.frame(psm, gpa, lti), !is.na(psm) & !is.na(gpa) & !is.na(lti))
pcorr.psm.gpa <- pcor.test(cordat2$psm, cordat2$gpa, cordat2$lti)
pcorr.psm.gpa
```

**Confidence intervals:** `r corr.ci(alpha = .05, corr = pcorr.psm.gpa$estimate, s = 1, n = pcorr.psm.gpa$n)`

#### Persistence and GPA:
```{r echo=FALSE}
cordat3 <- subset(data.frame(persist, gpa, lti), !is.na(persist) & !is.na(gpa) & !is.na(lti))
pcorr.persist.gpa <- pcor.test(cordat3$persist, cordat3$gpa, cordat3$lti)
pcorr.persist.gpa
```

**Confidence intervals:** `r corr.ci(alpha = .05, corr = pcorr.persist.gpa$estimate, s = 1, n = pcorr.persist.gpa$n)`

### Gender analyses

#### Persistence analyses

**Differences in persistence by gender:**
```{r echo=FALSE}
var.test(persist ~ gender)
t.test(persist ~ gender, var.equal=T)
describeBy(persist, gender)
```

**Perceptions of mobility predicting academic persistence, controlling for gender:**
```{r echo=FALSE}
summary(lm.beta(psm.persist.gender <- lm(persist ~ psm_c * gender_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
persist.regress <- lm(persist ~ psm_c * gender_c)
persist.regress.ci <- data.frame(round(confint(persist.regress, level = 0.95), digits=2))
colnames(persist.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
persist.regress.ci
```


#### GPA analyses

**Differences in GPA by gender:**
```{r echo=FALSE}
var.test(gpa ~ gender)
t.test(gpa ~ gender, var.equal=T)
describeBy(gpa, gender)
```

**Perceptions of mobility predicting GPA, controlling for gender:**
```{r echo=FALSE}
summary(lm.beta(psm.gpa.gender <- lm(gpa ~ psm_c * gender_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
gpa.regress <- lm(gpa ~ psm_c * gender_c)
gpa.regress.ci <- data.frame(round(confint(gpa.regress, level = 0.95), digits=2))
colnames(gpa.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
gpa.regress.ci
```


**Simple effects of gender on students with strong and weak perceptions of mobility:**
```{r echo=FALSE}
psm.gpa.lmres <- lmres(gpa ~ psm_c * gender_c, mydata)
summary(psm.gpa.ss <- simpleSlope(psm.gpa.lmres, pred="psm_c", mod1="gender_c"))
```

**CIs for female students:** `r round(psm.gpa.ss$simple_slope[1,1] + c(-1,1)*psm.gpa.ss$simple_slope[1,2]*qt(0.975, psm.gpa.ss$Df), digits=2)`  

**CIs for male students:** `r round(psm.gpa.ss$simple_slope[2,1] + c(-1,1)*psm.gpa.ss$simple_slope[2,2]*qt(0.975, psm.gpa.ss$Df), digits=2)`


**Academic persistence predicting GPA, controlling for gender:**
```{r echo=FALSE}
summary(lm.beta(persist.gpa.gender <- lm(gpa ~ persist_c * gender_c)))
```

**95% confidence intervals for unstandardized regression coefficients:**
```{r echo=FALSE}
gpapersist.regress <- lm(gpa ~ persist_c * gender_c)
gpapersist.regress.ci <- data.frame(round(confint(gpapersist.regress, level = 0.95), digits=2))
colnames(gpapersist.regress.ci) <- c("Lower (estimate)", "Upper (estimate)")
gpapersist.regress.ci
```


#### Mediation analysis

```{r include=FALSE}
persist.num <- as.numeric(persist_c)
psm.num <- as.numeric(psm_c)
summary(xm.model <- lm(persist ~ psm.num))
summary(full.model <- lm(gpa ~ psm.num + persist.num))
```

```{r echo=FALSE}
summary(bootstrap.inc <- mediate(xm.model, full.model, treat = "psm.num", mediator = "persist.num", sims = 5000, boot = TRUE, dropobs = TRUE))
```

**p-value of the indirect effect:** `r bootstrap.inc$d1.p`


```{r include=FALSE, warning=FALSE}
## Create and export tables ##

# full model sections
regtable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), round(coef(summary(modellist[[i]])), digits=3))
    assign(paste("ci", i, sep = ""), paste(round(get(paste("b", i, sep=""))[, 1], digits=2), " [", round(confint(modellist[[i]], level = 0.95)[,1], digits=2), ", ", round(confint(modellist[[i]], level = 0.95)[,2], digits=2), "]", sep = ""))
    assign(paste("t", i, sep = ""), round(summary(modellist[[i]])$coefficients[,3], digits=2))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$df[2],each=1+length(labels(modellist[[1]]))))
    assign(paste("p", i, sep = ""), round(summary(modellist[[i]])$coefficients[,4], digits=3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(all.vars(summary(modellist[[i]])$call)[1], ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rnames <- c("Intercept")
  for (j in 1:length(labels(modellist[[1]]))) {
    rnames <- append(rnames, labels(modellist[[1]])[j])
  }
  rownames(complete) <- rnames
  complete <- complete[-1,]
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=8)  # this can be an .rtf or a .doc
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models1 <- list(psm.persist.gender, psm.gpa.gender, persist.gpa.gender)
regtable(models1, "Table S5a (omnibus tests)")

# simple slope sections
sstable <- function(modellist, docname) {
  sections <- list(NA)
  colnam <- vector("list", length(modellist)*4)
  for (i in 1:length(modellist)) {
    assign(paste("b", i, sep = ""), round(summary(modellist[[i]])$simple_slope, digits=3))
    assign(paste("ci", i, sep = ""), paste(round(get(paste("b", i, sep=""))[, 1], digits=2), " [", c(round(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), digits=2)[1], round(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[i]]$Df), digits=2)[1]), ", ", c(round(get(paste("b", i, sep=""))[1,1] + c(-1,1)*get(paste("b", i, sep=""))[1,2]*qt(0.975, modellist[[i]]$Df), digits=2)[2], round(get(paste("b", i, sep=""))[2,1] + c(-1,1)*get(paste("b", i, sep=""))[2,2]*qt(0.975, modellist[[1]]$Df), digits=2)[2]), "]", sep = ""))
    assign(paste("t", i, sep = ""), round(get(paste("b", i, sep=""))[,3], digits=2))
    assign(paste("df", i, sep = ""), rep(summary(modellist[[i]])$Df,each=2))
    assign(paste("p", i, sep = ""), round(get(paste("b", i, sep=""))[,4], digits=3))
    assign(paste("section", i, sep = ""), rbind(cbind(get(paste("ci", i, sep = "")), get(paste("t", i, sep = "")), get(paste("df", i, sep = "")), get(paste("p", i, sep = "")))))
    sections[[i]] <- get(paste("section", i, sep = ""))
    colnam[1+(4*(i-1))] <- paste(modellist[[i]]$nomY, ": b (95% CIs)", sep="")
    colnam[2+(4*(i-1))] <- "t"
    colnam[3+(4*(i-1))] <- "df"
    colnam[4+(4*(i-1))] <- "p"
  }
  complete <- do.call(cbind, sections)
  colnames(complete) <- colnam
  rownames(complete) <- c(row.names(summary(modellist[[i]])$Points)[1], row.names(summary(modellist[[i]])$Points)[2])
  rtffile <- RTF(file=paste(docname, ".doc", sep=""), width=11, height=8.5, font.size=12)
  addTable(rtffile, cbind(rownames(complete), complete), col.justify="C", header.col.justify="C")
  done(rtffile)
}
models1.ss <- list(psm.gpa.ss)
sstable(models1.ss, "Table S5b (simple slopes)")
```
