---
title: "Scale Construction Study #1"
output:
  html_document:
    toc: true
    theme: default
---

## (1) Description of Data Files

```{r include=FALSE}
options(width=100)

library(RCurl)

# load data
csv <- getURL("https://raw.githubusercontent.com/abrowman/psm-jesp2017/master/Scale%20Construction%20Study%201/SocMobility_ScaleConstruction1.csv")
mydata <- read.csv(textConnection(csv))
```

The posted data for this study consists of 1 compiled, deidentified data file (total *N* recruited = `r nrow(mydata)`):

* `r nrow(subset(mydata, response==""))` participants excluded because they did not complete the task\
* `r nrow(subset(mydata, comments!="" & comments!="Did not complete task"))` participants excluded because they did not follow the instructions\

```{r include=FALSE}
mydata <- subset(mydata, response!="") # remove Ps who did not complete the task
mydata <- subset(mydata, comments=="") # remove Ps who did not follow the instructions
attach(mydata)
```

## (2) Analyses

**Total *N*:** `r nrow(mydata)`

```{r include=FALSE}
ses <- ifelse(education==1 | wealth==1 | occupation==1 | income==1 | class==1 | finances==1 | mobility==1 | possessions==1 | standard.of.living==1 | lifestyle==1 | affluence==1 | socioeconomic==1 | economic==1, 1, 0)

ses.tot <- rowSums(data.frame(education, wealth, occupation, income, class, finances, mobility, possessions, standard.of.living, lifestyle, affluence, socioeconomic, economic))

other.tot <- rowSums(data.frame(environment, regard.respect.reputation, ethnicity, friends, privilege, power, family, position.vs.others, good.person, importance, appearance, prestige, connections, success, wellbeing, social.conduct, authority, hierarchy, residence, influence, compassion, intelligence))
```

**Percentage of participants who gave a socioeconomic response:** `r round(sum(ses, na.rm=T)/nrow(mydata)*100, digits=1)`%

**T-test of whether Ps gave more socioeconomic or non-socioeconomic responses:**
```{r echo=FALSE}
t.test(ses.tot, other.tot, var.equal=T, paired=T)
```
\